Measuring the double-drift illusion with hand tracking

### 2021-03-01 Manuscript.Rmd
implemented half-way points correctly, updated re-tracing exp results / models

---
title: "Measuring the double-drift illusion with hand tracking"
author: 
- Bernard Marius 't Hart 1,*
- Denise Y.P. Henriques 1
- Patrick Cavanagh 1,2,3
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
bibliography: bibliography.bib
csl: journal-of-vision.csl
link-citations: yes
---

@@

```{r fig2, echo=FALSE, message=FALSE, fig.width=8, fig.height=5, fig.cap="\\label{fig:fig2}**Figure 2. Online Tracking Data** **Left:** Individual trajectories. Note that there are only a few obvious discontinuities (highlighted in red). In general, the traced paths showed rough linearity. **Top right:** Average heading in the middle 1 _s_ of each pass. Distance from origin is time, angle is angle. Dark blue lines the average heading, in light blue a polar heatmap with the distribution of heading. **Bottom right: The strength of the illusion in degrees angle deviation from the physical path over unity slope (internal / external velocity). In red are the average illusions strengths for 0 _cps_, ±1 _cps_ and ±3 _cps_, and in light blue the data for individual participants. Gray dashed line: unity. Black line: fit from Cavanagh & Tse (2019), 81% slope. Purple: fit to the data, 72% slope. **"}
plotBoundedTracking()
```

@@

We compare the strength of the illusion as measured with half-way points (see Fig 4, right), with an earlier study [@Cavanagh2019]. A line through the origin with slope 0.42 predicts illusion strength fairly well in this data, which is weaker than earlier (0.81) or with online tracking (0.72). We can also see that illusion strength measured with re-tracing is more variable. While the illusion is weaker, it is nevertheless there and depends on the relationship between internal and external speed, as expected.

@@

Where $L_y$ and $L_x$ are a temporal (y) and spatial (x) limit that are free parameters, and s corresponds to the calculated slope. If we fit these simple models by minimizing the mean square error (as the absolute distance in centimeters between predicted and actual reset points), the optimal parameters indicate a temporal limit of ~1.4 _s_ (AIC=-167.8) and a spatial limit of ~1.7 _cm_ (AIC=-208.6), as illustrated in Figure 6 (left and middle). This is close to the mean values of the respective coordinates (Figure 5, middle). The model with the spatial limit fits the data better than the model with a temporal limit ( _p_<.001, relative log-likelihood based on AICs). This again indicates that a spatial limit explains more of the data.

@@

The limits that optimally fit the model to the data are a 1.42 _cm_ spatial limit and a 0.28 _s_ temporal limit. While the AIC of the model (-209.8) is slightly lower than that of the model with the single spatial limit, it does not explain the data better ( _p_=.53). This means that a simple spatial limit accounts for most of the data. It is possible that the way we've modeled the additional temporal delay after crossing the spatial limit is not accurate, as it may not follow a normal distribution, which we do not account for here. However, regardless of whether or not a temporal limit adds anything, a spatial limit appears to explain the majority of spontaneous double-drift resets.

>>>>

---
title: "Measuring the double-drift illusion with hand tracking"
author: 
- Bernard Marius 't Hart 1,*
- Denise Y.P. Henriques 1
- Patrick Cavanagh 1,2,3
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
bibliography: bibliography.bib
csl: journal-of-vision.csl
link-citations: yes
---

@@

```{r fig2, echo=FALSE, message=FALSE, fig.width=8, fig.height=5, fig.cap="\\label{fig:fig2}**Figure 2. Online Tracking Data** **Left:** Individual trajectories. Note that there are only a few obvious discontinuities (highlighted in red). In general, the traced paths showed rough linearity. **Top right:** Average heading in the middle 1 _s_ of each pass. Distance from origin is time, angle is angle. Dark blue lines the average heading, in light blue a polar heatmap with the distribution of heading. **Bottom right:** The strength of the illusion in degrees angle deviation from the physical path over unity slope (internal / external velocity). In red are the average illusions strengths for 0 _cps_, ±1 _cps_ and ±3 _cps_, and in light blue the data for individual participants. Gray dashed line: unity. Black line: fit from Cavanagh & Tse (2019), 81% slope. Purple: fit to the data, 72% slope."}
plotBoundedTracking()
```

@@

We compare the strength of the illusion as measured with half-way points (see Fig 4, right), with an earlier study [@Cavanagh2019]. A line through the origin with slope 0.49 predicts illusion strength fairly well in this data, which is weaker than earlier (0.81) or with online tracking (0.72). We can also see that illusion strength measured with re-tracing is more variable. While the illusion is weaker, it is nevertheless there and depends on the relationship between internal and external speed, as expected.

@@

Where $L_y$ and $L_x$ are a temporal (y) and spatial (x) limit that are free parameters, and s corresponds to the calculated slope. If we fit these simple models by minimizing the mean square error (as the absolute distance in centimeters between predicted and actual reset points), the optimal parameters indicate a temporal limit of ~1.4 _s_ (AIC=-167.1) and a spatial limit of ~1.8 _cm_ (AIC=-220.7), as illustrated in Figure 6 (left and middle). This is close to the mean values of the respective coordinates (Figure 5, middle). The model with the spatial limit fits the data better than the model with a temporal limit ( _p_<.001, relative log-likelihood based on AICs). This again indicates that a spatial limit explains more of the data.

@@

The limits that optimally fit the model to the data are a 1.47 _cm_ spatial limit and a 0.33 _s_ temporal limit. While the AIC of the model (-226.9) is slightly lower than that of the model with the single spatial limit, it does not explain the data better ( _p_=.45). This means that a simple spatial limit accounts for most of the data. It is possible that the way we've modeled the additional temporal delay after crossing the spatial limit is not accurate, e.g. it may not follow a normal distribution. However, regardless of whether or not a temporal limit adds anything, a spatial limit appears to explain the majority of spontaneous double-drift resets.
